---
title: "`fbseqStudies` package installation and usage guide"
author: Will Landau
date: 2015
output: 
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
vignette: >
  \VignetteEngine{knitr::rmarkdown}
  \VignetteIndexEntry{`fbseqStudies` package installation and usage guide}
  \usepackage[utf8]{inputenc}
---

# Purpose

`fbseqStudies` runs a comparison study, comparing the `fbseq` package to other packages.

# System requirements

- Mac OS X or Linux. 
- R $\ge$ 3.2.0
- R packages `methods`, `parallel`, `snowfall`, `doMC`, `ggplot2`, `pracma`, `plyr`, `reshape2`, and `knitr`, available though the [Comprehensive R Archive Network (CRAN](https://cran.r-project.org/).
- The R package `edgeR`, available through [Bioconductor](http://bioconductor.org/). 
- R package `rstan`, the installation of which is explained on a [GitHub wiki](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started).
- A CUDA-enabled NVIDIA graphics processing unit (GPU) with compute capability 2.0 or greater.
- CUDA version 6.0 or greater.
- The `fbseq` and `fbseqCUDA` packages, available on GitHub.

More information about CUDA is available through [NVIDIA](http://www.nvidia.com/object/cuda_home_new.html). If your computer does not have these features, you can rent GPU time from [cloud computing enterprises](http://www.nvidia.com/object/gpu-cloud-computing-services.html) such as [Amazon Web Services](http://aws.amazon.com/ec2/instance-types/). See the installation vignettes for the `fbseq` and `fbseqCUDA` packages for installation details about those packages.

# Installation

## Option 1: install directly from GitHub.

For this option, you need the `devtools` packages, available from CRAN. Just open R and run 

```{r, eval=F}
library(devtools)
install_github("wlandau/fbseqStudies")
```

## Option 2: install from the source.

Open a command line program such as Terminal in Mac/Linux and enter the following commands.

```
git clone git@github.com:wlandau/fbseqStudies.git
R CMD build fbseqStudies
R CMD INSTALL fbseqStudies_0.0.tar.gz
```


Note: in `R CMD INSTALL fbseqStudies_0.0.tar.gz`, you may have to replace `fbseqStudies_0.0.tar.gz` by the name of whatever tarball comes out of `R CMD build`.

# Run the studies

## Everything in two lines.

Just open an R session and run 

```{r, eval = F}
library(fbseqStudies)
fbseqStudies()
```

## Breaking down the comparison study

You can run the steps of `fbseqStudies()` separately. First, create the simulation objects with `simulations()`. Output files are saved. These are lists $-$ each with a `fbseq::Scenario` object with the count data, design matrix, etc. $-$ and a list to store analyses of that scenario.

```{r, eval = F}
simulations()
```


# Appendix: setting up a CUDA-capable virtual server on Amazon Web Services

You can install this package and all its prerequisites on an Ubuntu 14.04 [g2.2xlarge instance](http://aws.amazon.com/ec2/instance-types/), which has a GPU. The following script configures the instance and installs all the prerequisites except for the `fbseq` and `fbseqCUDA` packages. After running the script, be sure to reboot the instance so that the CUDA drivers are loaded properly. 

```
echo "export PS1='\`hostname\` \`echo \$PWD\`> '" >> ~/.bashrc
echo "export PATH=/usr/local/bin:\$PATH" >> ~/.bashrc
source ~/.bashrc
sudo apt-get update

# There is a prompt here, and it should be the last interruption.
sudo apt-get -y install linux-image-generic

sudo apt-get -y install emacs24 git texlive-latex-base texinfo fonts-inconsolata r-base-core libxml2-dev r-cran-rgl linux-headers-3.13.0-57-generic openssl libcurl4-openssl-dev
sudo apt-get -y build-dep r-cran-rgl

sudo apt-get -y remove r-base-core
sudo wget http://cran.r-project.org/src/base/R-3/R-3.2.0.tar.gz
tar -xf R-3.2.0.tar.gz
cd R-3.2.0
./configure
make
sudo make install
cd
sudo rm -rf R-3.2.0*

echo 'library(utils); local({r <- getOption("repos"); r["CRAN"] <- "http://mirror.las.iastate.edu/CRAN/"; options(repos=r)}); assignInNamespace("q", function(save = "no", status = 0, runLast = TRUE){.Internal(quit(save, status, runLast))}, "base")' > .Rprofile

sudo Rscript -e 'install.packages(c("methods", "devtools", "parallel", "plyr", "ggplot2", "reshape2", "knitr", "snow", "snowfall",  "pracma", "coda", "doMC", "mclust", "Iso")); source("http://bioconductor.org/biocLite.R"); biocLite("edgeR", ask=FALSE)'

sudo Rscript -e 'source("http://mc-stan.org/rstan/install.R", echo = TRUE, max.deparse.length = 2000); install_rstan()'

# check https://developer.nvidia.com/cuda-downloads
# for the correct .deb file.
sudo apt-add-repository multiverse
sudo wget http://developer.download.nvidia.com/compute/cuda/repos/ubuntu1404/x86_64/cuda-repo-ubuntu1404_7.0-28_amd64.deb
sudo dpkg -i cuda-repo-ubuntu1404_7.0-28_amd64.deb
sudo apt-get update
sudo apt-get -y install cuda nvidia-cuda-toolkit nvidia-modprobe
sudo rm *.deb

echo "export PATH=/usr/local/cuda-7.0/bin:\$PATH" >> ~/.bashrc
echo "export LD_LIBRARY_PATH=/usr/local/cuda-7.0/lib64:\$LD_LIBRARY_PATH" >> ~/.bashrc
source .bashrc

cd /usr/local/cuda-7.0/samples/1_Utilities/deviceQuery
sudo make
cd

# Reboot and then run /usr/local/cuda-7.0/samples/1_Utilities/deviceQuery/deviceQuery
# to check GPUs.
sudo reboot

# check the GPUs
/usr/local/cuda-7.0/samples/1_Utilities/deviceQuery/deviceQuery

# troubleshoot at http://tleyden.github.io/blog/2014/10/25/cuda-6-dot-5-on-aws-gpu-instance-running-ubuntu-14-dot-04/

```
